Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-10-20 paper/main.tex
include citations from christian
Co-authored-by: Christian Fleck <christian.fleck@fdm.uni-freiburg.de>

Despite their popularity, \acp{abm} face multiple challenges in practise.
The construction of a fully functional \ac{abm} simulation from the ground up constitutes an
intricate and laborious process, necessitating a comprehensive array of skills. It is therefore
common practise to reuse existing simulation tools, which, in turn, brings along a number of
other problems.
Even if the existing tool is designed to be generalizable, it often comes with a pre-defined set
of parameters that need to be specified.
Furthermore, as we have discussed in our previous work~\cite{Pleyer2023}, these tools were most
often not designed to be flexible but rather built for specific purposes thus limiting their
generalizability \textbf{TODO CITATION}. Typically researchers estimate the parameter of their model
using methods such as Maximum Likelihood or Maximum a Posteriori.

[Gabor, Attila, and Julio R. Banga. ‘Robust and Efficient Parameter Estimation in Dynamic Models of Biological Systems’. BMC Systems Biology 9, no. 1 (2015): 74. https://doi.org/10.1186/s12918-015-0219-2., 
Banga, Julio R, and Eva Balsa-Canto. ‘Parameter Estimation and Optimal Experimental Design’. Essays in Biochemistry 45 (2008): 195–209. https://doi.org/10.1042/BSE0450195., 
Ashyraliyev, Maksat, Yves Fomekong-Nanfack, Jaap A Kaandorp, and Joke G Blom. ‘Systems Biology: Parameter Estimation for Biochemical Models’. FEBS Journal 276, no. 4 (2009): 886–902. https://doi.org/10.1111/j.1742-4658.2008.06844.x.] \\

check for parameter identifiability using methods like profile-likelihood or apply model reduction techniques \\

[Kreutz, Clemens, Andreas Raue, Daniel Kaschek, and Jens Timmer. ‘Profile Likelihood in Systems Biology’. The FEBS Journal, May 2013, n/a-n/a. https://doi.org/10.1111/febs.12276.
Raue, Andreas, Johan Karlsson, Maria Pia Saccomani, Mats Jirstrand, and Jens Timmer. ‘Comparison of Approaches for Parameter Identifiability Analysis of Biological Systems.’ Bioinformatics (Oxford, England), ahead of print, January 2014. https://doi.org/10.1093/bioinformatics/btu006.
Simpson, Matthew J. ‘Data-Informed Model Reduction for Inference and Prediction from Non-Identifiable Models’. Preprint, Synthetic Biology, 21 March 2025. https://doi.org/10.1101/2025.03.20.644462.]\\

>>>>

Despite their popularity, \acp{abm} face multiple challenges in practise.
The construction of a fully functional \ac{abm} simulation from the ground up constitutes an
intricate and laborious process, necessitating a comprehensive array of skills. It is therefore
common practise to reuse existing simulation tools, which, in turn, brings along a number of
other problems.
Even if the existing tool is designed to be generalizable, it often comes with a pre-defined set
of parameters that need to be specified.
Furthermore, as we have discussed in our previous work~\cite{Pleyer2023}, these tools were most
often not designed to be flexible but rather built for specific purposes thus limiting their
generalizability \textbf{TODO CITATION}.
Typically researchers estimate the parameter of their model using methods such as Maximum Likelihood
or Maximum a Posteriori~\cite{Gbor2015,Banga2008,Ashyraliyev2009}, check for parameter
identifiability using methods like profile-likelihood or apply model reduction
techniques~\cite{Kreutz2013,Raue2014,Simpson2025}.
